Global adjoint tomography -- Model GLAD-M25 \\ Supplementary Material

### 2020-02-14 main.tex
add psf figures
Updates from Overleaf

\subsection{Point-spread function analysis}

Checker-board tests or tests involving a chosen known structure, e.g., a plume or slab,
are frequently used to assess the quality of a tomographic inversion,
despite the fact that such tests have limited relevance~\citep[e.g.,][]{Leveque1993}.
Such tests are computationally expensive in FWI and adjoint tomography, because they
require roughly the same computational resources as the actual inversion.
In lieu of such tests, we previously performed two point-spread function (PSF) analyses~\citep{fichtner2011resolution}
for model GLAD-M15~\citep{bozdaug2016global}.
Here we repeat one of those tests (underneath Yellowstone) to show the improved resolution in GLAD-M25 relative to GLAD-M15,
and we perform three additional PSF tests: in the lower mantle underneath Afar, in the lower mantle beneath the Galapagos, and below the transition zone in Indonesia.

@@

Fig.~\ref{fig:??} shows the results of the four PSF tests we performed for model GLAD-M25.
The first test repeats an analysis performed by~\cite{bozdaug2016global} for GLAD-M15 in the upper mantle underneath Yellowstone.
That test was based on 253 earthquakes,
and Figs.~\ref{fig:??}a and~\ref{fig:??}b show our results for the same location based on half our dataset (720 events) and the full dataset (1,480 events), respectively.
We observe that the 75~km diameter anomaly is well recovered, with little tradeoff.
Figs.~\ref{fig:??}c--f shows the results of PSF tests in the lowermost mantle at the base of the African superplume,
in the lower mantle beneath the Galapagos,
and in the mid mantle below Indonesia.
In each case the Gaussian anomaly is well recovered, with minimal tradeoff with other model parameters.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/events_360.pdf}
  \caption{\small{Summary of the held-out database containing 360 earthquakes. Theses events were not used in the structural inversion. (a) Distribution of events color coded by depth range as in Fig.~\ref{fig:eventsstations}(a). (b) \& (c) Histograms of moment magnitude and depth. \tojeroen{Do I need to update this figure, to remove subfigure b and c?} }}
  \label{fig:events_360}
\end{figure}

>>>>

\subsection{Point-spread function analyses}

Checker-board tests or tests involving a chosen known structure, e.g., a plume or slab,
are frequently used to assess the quality of a tomographic inversion,
despite the fact that such tests have limited relevance~\citep[e.g.,][]{Leveque1993}.
Such tests are computationally expensive in FWI and adjoint tomography, because they
require roughly the same computational resources as the actual inversion.
In lieu of such tests, we previously performed two point-spread function (PSF) analyses~\citep{fichtner2011resolution}
for model GLAD-M15~\citep{bozdaug2016global}.
Here we repeat one of those tests (underneath Yellowstone) to show the improved resolution in GLAD-M25 relative to GLAD-M15,
and we perform three additional PSF tests: in the lower mantle underneath Afar, in the lower mantle beneath Easter Island, and below the transition zone in Sumatra.
Two additional PSF tests for anomalies below the US may be found in the SOM.

@@

Fig.~\ref{fig:psf_tests} shows the results of the four PSF tests we performed for model GLAD-M25.
The first test repeats an analysis performed by~\cite{bozdaug2016global} for GLAD-M15 in the upper mantle underneath Yellowstone.
That test was based on 253 earthquakes,
and Figs.~\ref{fig:psf_tests}a shows our result for the same location based on the current dataset of 1,480 events.
We observe that the 100~km diameter anomaly at 300~km depth is well recovered, with little tradeoff between~$\beta_\mathrm{V}$ and~$\beta_\mathrm{H}$ and~$c$ (See SOM).
Figs.~\ref{fig:psf_tests}b--d shows the results of PSF tests in the lower mantle beneath Afar,
in the lowermost mantle beneath Easter Island,
and in the mid mantle below Sumatra.
In each case the Gaussian anomaly is well recovered, with minimal tradeoff with other model parameters (See SOM).

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/psf_tests.pdf}
  \caption{\small{PSF tests.
  Each of the four panels labeled (a)--(d) shows a map view (Top Left) and cross section (Top Right) through a Gaussian perturbation, $\mathsf{\Delta}\mathsf{m}$\,, in~$\beta_\mathrm{V}$\,,
  and a map view (Bottom Left) and cross section (Bottom Right) through the recovered perturbation, $\mathsf{H}\cdot\mathsf{\Delta}\mathsf{m}$\,, in~$\beta_\mathrm{V}$.
  (a) 100~km diameter negative Gaussian anomaly at a depth of 300~km beneath Yellowstone.
  (b) 400~km diameter negative Gaussian anomaly at a depth of 2,000~km beneath Afar.
  (c) 300~km diameter negative Gaussian anomaly at a depth of 2,500~km beneath Easter Island.
  (d) 200~km diameter positive Gaussian anomaly at a depth of 900~km beneath Sumatra.
  }}
  \label{fig:psf_tests}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/events_360.pdf}
  \caption{\small{Held-out database containing 360 earthquakes. These events were not used in the structural inversion and are color coded by depth range as in Fig.~\ref{fig:eventsstations}(a).}}
  \label{fig:events_360}
\end{figure}

### 2020-02-14 supplementary_material.tex
Updates from Overleaf

>>>>

\section{Point Spread Function tests}

This section summarizes the complete Point Spread Function (PSF) tests conducted as part of this study.
Figs.~\ref{fig:psf_yellowstone}--\ref{fig:psf_su} show the results of PSF tests underneath Yellowstone, Easter Island, Afar, and Sumatra.
Each figure shows the original Gaussian perturbation, $\mathsf{\Delta}\mathsf{m}$\,, in map view and cross section, as well as the action of the Hessian on this model perturbation, $\mathsf{H}\mathsf{\Delta}\mathsf{m}$\,.
\begin{figure}
  \centering
  \includegraphics[width=0.65\textwidth]{figures/psf/yellowstone.pdf}
  \caption{\small{PSF test for Yellowstone.
  A 100~km diameter negative Gaussian anomaly is placed at a depth of 300~km beneath Yellowstone.
  Top row: Map view (Left) and cross section (Right) through a Gaussian perturbation in~$\beta_\mathrm{V}$.
  Second row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{V}$.
  Third row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{H}$.
  Bottom row: Map view (Left) and cross section (Right) through the recovered perturbation in the bulk sound speed~$c$. We conclude that there is little tradeoff between~$\beta_\mathrm{V}$ and~$\beta_\mathrm{H}$ and ~$c$.
  }}
  \label{fig:psf_yellowstone}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.65\textwidth]{figures/psf/easter.pdf}
  \caption{\small{PSF test for Easter Island.
  A 300~km diameter negative Gaussian anomaly is placed at a depth of 2,500~km beneath Easter Island.
  Top row: Map view (Left) and cross section (Right) through a Gaussian perturbation in~$\beta_\mathrm{V}$.
  Second row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{V}$.
  Third row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{H}$.
  Bottom row: Map view (Left) and cross section (Right) through the recovered perturbation in the bulk sound speed~$c$. We conclude that there is little tradeoff between~$\beta_\mathrm{V}$ and~$\beta_\mathrm{H}$ and ~$c$.  A 75~km diameter positive Gaussian anomaly is placed at a depth of 150~km beneath Yellowstone.
  Top row: Map view (Left) and cross section (Right) through a Gaussian perturbation in~$\beta_\mathrm{V}$.
  Second row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{V}$.
  Third row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{H}$.
  Bottom row: Map view (Left) and cross section (Right) through the recovered perturbation in the bulk sound speed~$c$. We conclude that there is little tradeoff between~$\beta_\mathrm{V}$ and~$\beta_\mathrm{H}$ and ~$c$.}}
  \label{fig:psf_easter}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.65\textwidth]{figures/psf/afar.pdf}
  \caption{\small{PSF test for Afar.
    A 400~km diameter negative Gaussian anomaly is placed at a depth of 2,000~km beneath Afar.
  Top row: Map view (Left) and cross section (Right) through a Gaussian perturbation in~$\beta_\mathrm{V}$.
  Second row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{V}$.
  Third row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{H}$.
  Bottom row: Map view (Left) and cross section (Right) through the recovered perturbation in the bulk sound speed~$c$. We conclude that there is little tradeoff between~$\beta_\mathrm{V}$ and~$\beta_\mathrm{H}$ and ~$c$.
  }}
  \label{fig:psf_afar}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.65\textwidth]{figures/psf/su.pdf}
  \caption{\small{PSF test for Sumatra.
    A 200~km diameter positive Gaussian anomaly is placed at a depth of 900~km beneath Sumatra.
  Top row: Map view (Left) and cross section (Right) through a Gaussian perturbation in~$\beta_\mathrm{V}$.
  Second row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{V}$.
  Third row: Map view (Left) and cross section (Right) through the recovered perturbation in~$\beta_\mathrm{H}$.
  Bottom row: Map view (Left) and cross section (Right) through the recovered perturbation in the bulk sound speed~$c$. We conclude that there is little tradeoff between~$\beta_\mathrm{V}$ and~$\beta_\mathrm{H}$ and ~$c$.
  }}
  \label{fig:psf_su}
\end{figure}
